
\chapter{整合}

\section{错题分析}

6、 [单选] 项目经理正在帮助公司推出一件新产品，该产品与将于未来几年推出的一系列新产品很相似。在关闭当前项目时，项目经理应该做什么来帮助未来的项目？

The project manager is helping the company launch a new product that is similar to a series of new products to be launched in the coming years. What should the project manager do to help future projects when closing the current project?

A：向新项目经理提供项目绩效报告

Provide a project performance report to the new project manager

B：记录项目相关方提供的经验教训

Document the lessons learned by project stakeholders in providing relevant information

C：与项目相关方一起审查最后更新的问题日志

Review the last updated issue log together with project stakeholders

D：确保风险登记册处于最新状态，并让项目相关方参与进来

Ensure that the risk register is up-to-date, and involve the project stakeholders

正确答案：B

你的答案：C

解析：解析：PMBOK（6）121-4.7结束项目或阶段。“帮助未来的项目”，有类似的关键词时，要考虑到经验教训登记册和组织过程资产。又提到了“在关闭当前项目时”，在收尾过程中对经验教训的总结非常重要，也是常出的考点。

7、 [单选] 一场重要会议结束后，团队收到客户的负面反馈，原因在获得需求规范时所使用的方法上，团队才意识到这一点。项目经理应该做什么？

After an important meeting, the team receives negative feedback from the customer. The team attributes this issue to the approach used to obtain requirement specifications. What should the project manager do?

A：检查相关方参与计划

Check the stakeholder engagement plan

B：提供经验教训知识库

Provide a lessons learned repository

C：更新范围管理计划

Update the scope management plan

D：与团队一起核实验收标准

Verify the acceptance criteria with the team

正确答案：B

你的答案：D

解析：解析：PMBOK（6）P98-4.4管理项目知识。“获得需求规范时所使用的方法”，问题经发生了，尽量规避以后再次出现。选项A，问题的核心是用错了规范，而不是相关方的态度。选项C，情景里面没有强调是哪一次会议，如果是收尾过程的会议，范围管理计划的更新将失去意义。选项D，情景中没有提到验收标准存在问题，排除。

12、 [单选] 项目关键路径上的任务所需的设备交付延迟了两个星期。项目经理应该做什么？

The delivery of equipment required for tasks on the critical path of the project was delayed by two weeks. What should the project manager do?

A：在风险登记册中登记该事件，并立即通知项目发起人

Register the event in the risk register and notify the project sponsor immediately

B：在问题日志中解决该事件，并在下次项目状态会议上讨论

Resolve the event in the issue log and discuss it in the next project status meeting

C：在问题日志中记录该事件，并立即通知相关方

Document the event in the issue log and notify the stakeholder immediately

D：在风险登记册中记录该事件，并在下次项目状态会议上讨论

Document the event in the risk register and discuss it in the next project status meeting

正确答案：B

你的答案：A

解析：解析：PMBOK（6）P96-4.3.3.4指导与管理项目工作-问题日志。问题已经出现，所以不应将其当成不确定性事件进行管理，排除AD。问题日志侧重问题的跟踪解决和聚焦，而不是简单的记录和上报，排除选项C。

16、 [单选] 在设计阶段进行的一项分析表明，项目可能在财务方面不可行。项目经理应该审查什么？

An analysis carried out during the design phase indicated that the project might not be financially possible. What should the project manager review?

A：项目前五年的运营成本

Operating costs for the first five years of the project

B：基于预算线的非营运性投资成本

Non-operating investment cost based on the budget line

C：商业论证中的社会和业务目标效益

Social and business objective benefits in business case

D：基于之前类似投资的预期回报效益

Expected returns based on previous similar investments

正确答案：C

你的答案：B

解析：解析：PMBOK（6）P77-4.1.1.1商业论证。设计阶段发现财务方面不可行，要回溯看一下商业论证的结果，判断导致问题产生的原因。选项AB提到的成本是在规划后的，排除。选项D，属于商业论证中可能出现的内容。


17、 [单选] 在项目执行结束时，项目经理保留了一些团队成员。在这个时间点，团队应该主要关注什么？

At the end of the project execution, the project manager retains some team members. What should the team focus on at this point in time?

A：协商清理、现场恢复、调试和人员培训

Follow-up clean-up, on-site recovery, commissioning and personnel training

B：根据公司政策支持项目的行政收尾程序

Administrative closing procedures to support projects in accordance with company policy

C：运营新建成的设施，直到负责人完全有能力运营为止

Operation of the newly built facilities until then they are fully capable of operating

D：提供支持，作为与移交相关的保修期的组成部分

Supply of support as part of the warranty period associated with the transfer

正确答案：B

你的答案：A

解析：解析：PMBOK（6）121-4.7结束项目或阶段。本过程的主要作用是，存档项目或阶段信息，完成计划的工作，释放组织团队资源以展开新的工作。向运营团队移交、提供维保服务、培训属于收尾过程的活动，选项ACD可能出现在选项B中。


19、 [单选] 在项目审查会议期间，相关方A指出已更新一项可交付成果来反映来自相关方B的新需求。但是，相关方A未获告知并且不同意这一变更。项目经理下一步应该做什么？

During the project review session, Skateholder A indicated that it had updated a deliverable to reflect new requirements from Stakeholder B. However, Skateholder A was not informed and did not agree to the change. What should the project manager do next?

A：检查变更日志以确认相关方A的主张

Check the change log to validate the claim of stakeholder A

B：从项目中删除新需求

Remove new requirements from the project

C：要求相关方A和B之间召开会议以解决冲突

Request a meeting between stakeholders A and B to resolve the conflict

D：将新需求升级上报给项目发起人

Escalate new requirements to the project sponsor

正确答案：A

你的答案：C

解析：解析：PMBOK（6）P120，4.6.3.3项目文件更新-变更日志。变更日志用于记录项目期间发生的变更。当前相关方对变更请求存在争议，首先查看变更日志中的审批记录。

20、 [单选] 在完成一个项目的规划阶段之后，但在执行之前，项目经理需要沟通并获得对项目目标的承诺。项目经理下一步应该做什么？

After completing the planning phase of a project, but before it can be implemented, the project manager needs to communicate and gain commitment to the project objectives. What should the project manager do next?

A：与所有相关方一起举行开工会议

Hold an kick-off meeting with all stakeholders

B：通过电子邮件将项目管理计划发送给所有相关方

E-mail the project management plan to all stakeholders

C：与每个相关方通话，以解释他们的角色与职责

Call each stakeholder to explain their roles and responsibilities

D：要求项目发起人与所有相关方沟通该计划

Request the project sponsor to communicate the plan to all stakeholders

正确答案：A

你的答案：D

解析：PMBOK 86 4.2.2.4 制定项目管理计划—工具与技术—会议。题目关键“规划后，执行前，获得相关方承诺“，开工会议（kick-off/start up) "旨在传达项目目标，获得团队对项目的承诺以及阐明相关方的角色和职责“，选A

24、 [单选] 在项目启动阶段，项目经理发现，事业环境因素中最近的变更已对预算产生影响，导致团队完成全部工作范围的事项变得不切实际。项目经理应如何缓解这种情况？

During the project initiation phase, the project manager finds that recent changes of the enterprise environmental factors have had an impact on the budget, making it impractical for the team to complete the full scope of work. How should the project manager deal with this situation?

A：根据纠正后的预算制定项目管理计划

Develop the project management plan according to the corrected budget

B：联系项目发起人，更正项目章程并寻求其重新予以批准

Contact the project sponsor to correct the project charter and seek re-approval

C：找机会缩小项目范围，使之与获批的预算一致

Look for opportunities to narrow the project scope to align with the approved budget

D：将范围方面的顾虑列为一种风险，并制定应急计划

Identify concerns on project scope as a risk and develop contingency plans

正确答案：B

你的答案：A

解析：PMBOK P81 4.1.3.1 项目章程。关键词“启动阶段“，事业环境因素变更影响了工作范围，因为在启动阶段，只有项目章程，含有高层级的需求等信息，章程由发起人负责，因此选B，对项目章程更新并再次获得批准。此阶段不存在项目管理计划、范围，ACD错

25、 [单选] 项目经理正在管理一个项目，其供应商负责采购最终组装所需的一种重要材料，但在进入组装阶段时却发现这种重要材料的交货跟踪记录不翼而飞，而且这种材料尚未装运。如果这种材料未按估算的交货时间运达，这可能会导致整个进度计划发生延迟。为解决该问题，项目经理接下来该做什么？

The project manager was managing a project. The vendor was responsible for purchasing an important material required for final assembly, but it was discovered that the delivery track record of this important material was missing, and this material had not yet been shipped during the assembly phase. If this material does not arrive on estimated delivery time, it may cause a delay in the overall schedule. What should the project manager do next to solve this issue?

A：记录该情况的所有详细信息，并更新问题日志

Document all the details of the situation and update the issue log

B：将该情况报告给项目发起人，并查明延长项目时限将会产生什么样的影响

Report the situation to the project sponsor and identify what the impact of a time extension will be

C：转向先前实施的评估过程中发现的次优的合格采购供应商

Move to sub-optimal qualified procurement vendors identified in the previously implemented evaluation process

D：与采购团队一起评估该情况，并探索加快交货的所有可能的解决方案

Work with the procurement team to assess the situation and explore all possible solutions to expedite delivery

正确答案：A

你的答案：D

解析：PMBOK P96 4.3.3.3 问题日志。根据题意，材料跟踪记录丢失，已经发生了，属于问题，首先应该记入问题日志，其次分析影响等具体行动。因此选A。

27、 [单选] 在向相关方提交项目管理计划时,项目经理被要求在项目管理计划内记录现有效益管理计划中已确定的项目效益。项目经理该做什么？

When submitting a project management plan to the stakeholder, the project manager is required to document the project benefits identified in the existing benefits management plan within the project management plan. What should the project manager do?

A：确认已识别的项目效益会成为规划的一部分

Validate that identified project benefits will be part of the plan

B：确认在执行阶段会制定效益管理计划

Validate that a benefits management plan will be developed during the execution phase

C：确认已识别的项目效益会受到适当监督

Validate that identified project benefits will be properly monitored

D：确认在项目收尾阶段会制定效益管理计划

Validate that a benefits management plan will be developed at the closure phase of the project

正确答案：A

你的答案：C

解析：解析：PMBOK（6）P86-4.2.3.1项目管理计划。 题干关键词“向相关方提交项目管理计划时”。 项目管理计划提交相关方批准时，基准尚未被批准，此时可以对项目管理计划进行多次更新，并无需遵循正式流程，故选A; BCD在规划阶段之后的执行、监控、收尾阶段，再对项目管理计划变更，需要遵循整体变更控制流程。

29、 [单选] 在项目执行期间，一位技术专家必须暂时休假。在其暂时休假期间，必须开发一项重要的可交付成果。项目经理没有执行这个活动的备用资源。若要在未来的项目中预防这种情况，项目经理应该做什么？

During the execution of the project, a technical specialist must be on leave for the duration of the project. During his sabbatical, he must develop an important deliverable. The project manager does not have the spare resources to execute this activity. What should the project manager do to prevent this in future projects?

A：将该问题记录在风险登记册中并对其进行优先级排序

Document the issue in the risk register and prioritize it

B：修订可交付成果的截止日期

Revise the deadline for deliverables

C：更新经验教训知识库

Update the lessons learned repository

D：在项目进度计划中包含缓冲时间

Include buffer time in the project schedule

正确答案：C

你的答案：D

解析：解析：PMBOK（6）P104-4.4.3.1经验教训登记册。 题干关键词“在未来项目中预防这种情况”。 为了给未来项目提供经验教训，项目经理应该把相关信息归入经验教训知识库，成为组织资产的一部分，故选C。 选项ABD都没有直接点出知识库的更新，错误。

