
\chapter{沟通}

\section{错题}

1、 [单选] 某大型项目的项目经理观察到，相关方的沟通方式将危及项目的成功。项目经理应该做什么来应对这种情况？

The project manager for a large project observes that the way the stakeholder communicates would jeopardize the success of the project. What should the project manager do to deal with this situation?

A：将该问题上报给发起人，以确保这种情况不会重复发生

Escalate the issue to the sponsor to ensure that it does not recur

B：参照沟通管理计划，寻求让相关方了解相关情况方面的指导

Refer to the communications management plan and seek guidance for stakeholders to understand the relevant situation

C：审查事业环境因素(EEFs)，以建立适当的沟通渠道

Review Enterprise Environmental Factors (EEFs) to establish appropriate communication channels

D：直接与相关方讨论该问题，以纠正沟通问题

Discuss the issue directly with the stakeholder to correct the communication problem

正确答案：D

你的答案：B

解析：解析：PMBOK（6）P388-10.3监督沟通。监督沟通是确保满足项目及其相关方的信息需求的过程。“沟通方式”影响到了项目的成功，要对沟通的问题进行分析，根据分析的结果做出调整。选项A，是评估后可能做出的决策。选项BC，可能出现在D选项的评估过程。



6、 [单选] 某项目已落后于进度计划，项目经理已通过最新的状态报告将此问题上报给指导委员会，某职能部门交付经理建议，他们尚不清楚状态更新中的报告的进度计划延迟情况。项目经理该做什么？

A project is behind schedule, the project manager has escalated the issue to the steering committee through the latest status report, and a functional delivery manager claims that they are not yet aware of the schedule delays in the status update report. What should the project manager do?

A：让职能部门经理参阅定期团队状态更新，该问题会在上报前首先予以报告

Have the functional manager refer to regular team status updates, and the issue will be escalated before escalation

B：更新沟通管理计划，以反映职能部门经理的需求，并发送状态更新的早期草案

Update the communications management plan to reflect the requirements of the functional manager and send early draft status updates

C：向指导委员会全体成员发送每周进展情况报告，并在开会前收集反馈

Send weekly progress reports to all members of the Steering Committee and gather feedback before meetings

D：查阅相关方参与计划，并执行相关活动，以确保符合所有相关方的沟通需要

Review the stakeholder engagement plan and execute related activities to ensure that the communication requirements of all stakeholders are met

正确答案：D

你的答案：B

解析：解析：PMBOK（6）P388，10.3-监督沟通。虽然报告已发送给相关方，但相关方仍然对需要沟通的信息不了解，说明沟通管理计划存在缺陷。要想解决问题，需要从相关方的沟通需求开始梳理，更新沟通管理计划。选项A，没有从源头上解决问题。选项B，在选项D之后。选项C，之前发送过。


7、 [单选] 一个项目相关方抱怨收到太多信息，另一个相关方却抱怨收到的信息太少，项目经理应该做什么？

One project stakeholder complains about receiving too much information, while another complains about receiving too little information. What should the project manager do?

A：标记报告的不同部分，并要求相关方仅关注属于他们的部分

Mark the different parts of the report and ask the stakeholders to focus only on the parts that belong to them

B：制定一份执行、负责、咨询和知情（RACI）图，以确保满足相关方的沟通需求

Develop an Responsible, Accountable, Consulted and Informed (RACI) diagram to ensure that the communication requirements of stakeholders are met

C：安排召开两个相关方之间的会议以达成妥协

Arrange a meeting between two stakeholders to reach a compromise

D：进行沟通需求分析以制定一份整体沟通管理计划

Conduct communication requirements analysis to develop an overall communications management plan

正确答案：D

你的答案：B

解析：解析：PMBOK（6）P388，10.3-监督沟通。对于沟通管理计划的更新，要根据相关方对信息获取的需求进行分析后决策，而不能仅仅根据相关方的“抱怨”冒然更新，所以选项A错误。选项B，RACI通常用于分配职责与角色。选项C，问题的焦点不是冲突，而是信息传递上存在缺陷。


8、 [单选] 在客户举行的一次会议上，项目经理发现了一个与可交付成果质量相关的重要问题。项目经理该使用哪些沟通方法？

At a meeting with the customer, the project manager identified an important issue related to the quality of deliverables. What communication methods should project manager use?

A：正式和口头

Formal and oral

B：正式和书面

Formal and written

C：非正式和书面

Informal and written

D：非官方和书面

Unofficial and written

正确答案：B

你的答案：A

解析：“可交付成果质量相关的重要问题”，关键信息的传递，需要采用正式、书面的沟通方法。


9、 [单选] 刚刚加入一个项目的项目经理注意到团队成员已与客户建立了电子邮件沟通关系。 项目经理担心对项目信息的共享方式缺乏控制。项目经理应该做什么？

The project manager who has just joined a project notices that the team members have established an email communication relationship with the customer. The project manager is worried about the lack of control over how project information is shared. What should the project manager do?

A：确保所有客户交互与批准的沟通管理计划保持一致

Ensure that all customer interactions are consistent with the approved communications management plan

B：告知团队所有沟通必须通过项目经理

Inform the team that all communication must go through the project manager

C：要求所有未来的沟通都包含项目发起人

Require all future communications to include the project sponsor

D：审查相关方参与计划，以确定沟通线路的权限

Review stakeholder engagement plan in the plan to determine the authority of communication lines

正确答案：A

你的答案：D

解析：解析：PMBOK（6）P388，10.3-监督沟通。注意审题，”担心对项目信息的共享方式缺乏控制“，说明担心的是沟通管理失控，最好的办法就是要求参与沟通的相关方严格按照沟通管理计划的约定进行沟通。选项BC，必须经过项目经理或项目发起人错误，除非在沟通管理计划中明确。选项D，情景中关注的焦点不是新出现的沟通渠道是否合理。



10、 [单选] 项目经理开展的一个项目的相关方位于不同的国家/地区，各自讲着不同的当地语言，项目经理应该如何减少错误传达的发生？

The stakeholders in a project carried out by the project manager are located in different countries/regions and each speak a different local language. How should the project manager reduce the occurrence of miscommunication?

A：让相关方来到项目经理所在国家/地区

Bring the stakeholders to the project manager's country/region

B：使用商定的共同语言

Use the agreed common language

C：仅使用书面沟通形式

Use written communication only

D：为相关方安排翻译人员

Arrange translators for the stakeholders

正确答案：B

你的答案：D

解析：解析：PMBOK（6）P366，10.1-规划沟通管理。相关方分散于不同地区，且使用不同语言。在规划沟通管理时要针对这个痛点去解决，采用共同的语言是一种常用的解决办法。选项A，让所有的相关方聚集于同一地区不现实。选项CD，书面沟通和安排翻译人员也是建立在采用商定的共同语言基础上。


14、 [单选] 项目经理为一个项目识别了相关方。但是，在分发更新文件时，三位相关方却未收到更新。应该采取哪一项措施来避免这种情况？

A project manager identified stakeholders for a project. However, three stakeholders failed to receive an update when it was distributed. What should have been done to avoid this situation?

A：遵循沟通管理计划

Follow the communications management plan

B：遵循相关方参与计划

Follow the stakeholder engagement plan

C：更新项目进度计划，以包含状态分发

Update the project schedule to include the status distribution

D：更新事业环境因素，以包含状态分发

Update the enterprise environmental factors to include the status distribution

正确答案：A

你的答案：B

解析：解析：PMBOK（6）P379，10.2-管理沟通。管理沟通是确保项目信息及时且恰当地收集、生成、发布、存储、检索、管理、监督和最终处置的过程。未收到文件，沟通管理出现了问题，如何”避免“，遵循沟通管理计划。选项B，相关方参与计划侧重于如何调动相关方参与到项目当中。选项CD，是在解决当前问题后做出的调整。


15、 [单选] 在一个正在进行的客户特定产品开发项目中，客户的项目经理发生变化，新项目经理没有收到关于项目可交付成果的更新信息。 项目经理应该怎么做？

During an ongoing, customer-specific product development project, the customer's project manager is changed. The new project manager fails to receive updates about the project's deliverables. What should the project manager do?

A：更新变更日志

Update the change log

B：查阅范围文件

Refer to the scope document

C：查阅干系人登记册

Refer to the stakeholder register

D：更新沟通管理计划

Update the communications management plan

正确答案：D

你的答案：C

解析：解析：PMBOK（6）P379，10.2-管理沟通。管理沟通是确保项目信息及时且恰当地收集、生成、发布、存储、检索、管理、监督和最终处置的过程。未收到信息，沟通管理出现了问题，首先查阅沟通管理计划，分析问题出现的原因，必要时根据相关方沟通需求更新沟通管理计划。选项AB，尚未提出变更请求。选项C，属于D选项的一个环节。


17、 [单选] 某虚拟团队正努力按时完成项目可交付成果。项目经理将举行会议，以便提供直接的解决方案。项目经理应做什么来组织有效会议？

A virtual team is making great efforts to complete the project deliverables on time. The project manager will hold a meeting so as to provide a direct solution. What should the project manager do to organize an effective meeting?

A：邀请所有项目相关方与会，以便他们了解项目延迟情况

Invite all project stakeholders and people so that they can understand the project delay

B：通过保存会议纪要来记录会议情况，并让团队成员以匿名方式发表意见

Keep a record of the request and ask team members to comment anonymously

C：向项目团队提供更新后的项目管理计划，该计划应反映他们在会前能够发现的延迟情况

Provide the project team with an updated project management plan that reflects the delays they can find before the meeting

D：会前发送进展报告，以便团队了解可交付成果对项目的影响

Send a report before the meeting so that the team understands the impact of deliverables on the project

正确答案：D

你的答案：C

解析：解析：PMBOK（6）P386，10.2.2.6人际关系与团队技能-会议管理。开会是为了解决问题，所以是否邀请所有的项目相关方参加会议，要根据会议的目标来安排，排除选项A。问题B，会议纪要在会议结束后输出。选项C，情景中没有提到延迟的风险。



18、 [单选] 某项目经理的团队成员位于三大洲。项目经理发现，某些项目团队成员无法及时收到重要信息。项目经理该做什么？

A project manager's team members are located on three continents. The project manager found that some project team members did not receive important information in a timely manner. What should the project manager do?

A：对所有进度计划做出调整，以便每个人都能同时工作。从而实现实时沟通和决策

Make adjustments to all schedules so that everyone can work at the same time. This enables real-time communication and decision-making

B：轮流安排电话会议和视频会议，以便顾及每个人的日程安排，使团队成员能够进行互动

Take turns scheduling conference calls and videocons to take into account each person's schedule and enable team members to interact

C：在分散工作团队之间仅使用电子邮件和书面通信，以便团队成员保留书面记录

Use only e-mail and written communications between decentralized work teams so that team members can keep written records

D：启用各团队内的一个中心人物，限制分散工作团队之间的沟通，从而减少错误传达的发生

Enable a central person within each team to limit communication between decentralized work teams, reducing the occurrence of false communication

正确答案：B

你的答案：D

解析：解析：PMBOK（6）P392-10.3.2.5会议。 题干关键词“无法及时收到重要信息”。 本题用排除法为宜。 选项A错误，关键词说明这是一个沟通问题，而不是进度计划出了问题； 选项B正确，交互式的沟通（会议）有利于团队成员交流和分享信息； 选项C错误，推式沟通效果不如交互式沟通； 选项D错误，为了减少错误传达而限制沟通是消极的做法。



