
\chapter{相关方}

\section{错题解析}

3、 [单选] 项目经理正在记录一个有大量相关方项目的相关方参与情况。项目经理对相关方彼此之间的关系不清楚。项目经理该做什么来更清楚地了解这些相关方？

The project manager is documenting the stakeholder engagement of a project involved with a large number of stakeholders. The project manager is not clear about the relationship between the stakeholders. What should the project manager do to understand these stakeholders more clearly?

A：构建相关方思维导图

Build a Mind-Mapping for the stakeholders

B：审查各相关方过往的项目文件

Review past project documents of all stakeholders

C：审查相关方参与度评估矩阵

Review the stakeholder engagement assessment matrix

D：请求与发起人和团队开会讨论相关方事宜

Request a meeting with the sponsor and team to discuss matters concerned with the stakeholders

正确答案：A

你的答案：C

解析：解析：PMBOK（6）P521，13.2.2.5规划相关方参与-思维导图。思维导图用于对相关方信息、相互关系以及他们与组织的关系进行可视化整理。选项C，相关方参与度评估矩阵用于将相关方当前参与水平与期望参与水平进行比较。选项D，发起人通常出现在启动和收尾两个过程。

10、 [单选] 项目经理计划召开开工会议，并准备一份全面的演示文稿，在开工会议期间出席会的人员承诺支持该项目，但是在项目开始后项目经理注意到一些相关方根本没有参与。若要避免这个问题，项目经理应该事先做什么？

The project manager plans to hold an kick-off meeting and prepare a comprehensive presentation in which the participants commit to supporting the project, but after the project starts, the project manager notices that some stakeholders are not involved at all. What should the project manager do beforehand to avoid this issue?

A：进行相关方分析

Execute a stakeholder analysis

B：确保相关方充分了解范围

Ensure that the stakeholders are fully aware of the scope

C：确保所有相关方均收到开工会议记录，其中记录了所有相关方协议

Ensure that all stakeholders receive the minutes of the initiation meeting, which document all stakeholder agreements

D：请求发起人确保相关方职能经理的承诺

Request the sponsor to ensure the commitment of the stakeholder functional manager

正确答案：A

你的答案：C

解析：解析：PMBOK（6）P533-13.1.2.3相关方分析。题干中的关键字：“项目开始后”“一些相关方没有参与”“事先做什么”。根据关键字，题目要求找出相关方知识领域在执行之前的相关内容，比如相关方参与计划或相关方登记册等，所以选项A正确。题目中明确相关方出席了开工会议，所以相关方了解范围也同时收到了会议记录，排除BC。发起人不负责确保职能经理承诺，排除D。

11、 [单选] 一个高收入且项目团队成员超过50人的大型软件开发项目正处于执行阶段，项目需求已收集，且项目团队在两个月内完成了第一模块，结果已提交给相关方，并且团队获得了批准，但是相关方通知团队新需求之前未沟通过。 若要避免这个问题，项目经理应该在执行阶段做什么？

A large software development project with a high-income team with more than 50 project members is in the execution phase, project requirements have been collected, and the project team completed the first module within two months, the results have been submitted to the stakeholder, and the team has been approved, but the stakeholder have not communicated before notifying the team of the new requirements. To avoid this issue, what should the project manager do during the execution phase?

A：召开状态会议和回顾会议，以发现意外的需求并更新项目管理计划

Hold status meetings and retrospective meetings to identify unexpected needs and update project management plans

B：与发起人交流并解释该情况，以获得更多时间来修改新的相关方需求

Talk to the sponsor and explain the situation to get more time to modify the new stakeholder requirements

C：举行头脑风暴会议以再次收集范围管理计划中的需求，并一起执行所有需求

Hold brainstorm meetings to re-collect the requirements in the scope management plan and execute all the requirements together

D：创建相关方分析，以了解个人期望，帮助规划正确的方法以便立即编译所有需求

Create stakeholder analysis to understand individual expectations and help plan the right approach to compile all requirements immediately

正确答案：D

你的答案：A

解析：解析：PMBOK（6）P533-13.3管理相关方参与。题干中的关键字：“相关方通知新需求之前未沟通”“若要避免这个问题”。根据题意需要寻找防止问题发生的方法，题目中明确相关方的需求未识别。所以需要进行相关方分析来确认相关方对项目的期望。选项D正确。回顾和获得更多的时间都时解决问题的方法，不是防止的方法，排除AB。头脑风暴也是收集需求的办法，但是不解决不了解相关方需求的问题，排除C。

12、 [单选] 在部署了一个企业资源管理系统期间，财务总监辞职并由新财务总监取代。项目经理应该做什么？

During the deployment of an enterprise resource management system, the Finance Director resigned and was replaced by the new Finance Director. What should the project manager do?

A：发布更新联系人列表和相关方登记册的变更请求，以包含该财务总监

Post a change request to update the contact list and stakeholder register to include the Finance Director

B：将项目文档发送给该财务总监，并请求其进行审查

Send the project document to the Finance Director and request his or her review

C：请新财务总监更新项目成本管理计划

Request the new Finance Director to update the project cost management plan

D：与新的财务总监开会，以介绍项目并了解它们的参与情况

Have a meeting with the new Finance Director to present the projects and understand their engagement

正确答案：A

你的答案：D

解析：解析：PMBOK（6）P512-13.1.3.1，相关方登记册。题干中的关键字：“财务总监辞职并由新财务总监取代”。根据题目可知，有新的相关方加入，需要更新相关方登记册，选项A正确。财务总监不一定有审查和更新项目文档的权限，排除BC。了解新的财务总监的需求在更新相关方登记册之后，排除D。

13、 [单选] 一位项目经理是在执行阶段中间刚刚任命的，该项目被延迟，且两个重要相关方对项目目标有分歧，文档规定不能与任意相关方直接互动，应将此角色留给公司的首席信息官。项目经理应该做什么？

A project manager was appointed in the middle of the execution phase, the project was delayed, and two key stakeholders disagreed on the project objectives, and the documentation stated that no direct interaction with any of the stakeholders should be left to the company's Chief Information Officer. What should the project manager do?

A：执行根本原因分析并升级上报给这两个重要的相关方

Execute root cause analysis and escalate to these two important stakeholders

B：执行相关方分析，并将该问题升级上报给高级管理层

Execute a stakeholder analysis and escalate the issue to senior management

C：执行假设条件和制约因素分析，并审查相关方参与计划

Execute an analysis of assumptions and constraints and review the stakeholder engagement plan

D：执行数据分析并审查相关方的角色与职责矩阵

Execute data analysis and review the roles and responsibilities matrix of the stakeholders

正确答案：B

你的答案：D

解析：解析：PMBOK（6）P512-13.1.2.3，相关方分析。题干中的关键字：“两个相关方对项目目标有分歧”“文档规定不能与任意相关方直接互动”。由题干可知，相关方对目标有分歧，需要了解每个相关方他们的目标时什么。所以执行分析，并根据提示提交给高层，选项B正确。根本原因分析没有相关方分析直接，排除A。相关方参与计划与项目目标无关，排除C。角色和职责矩阵与项目目标无关，排除D。

14、 [单选] 某有着众多相关方的复杂项目处于执行阶段，此时由于管理层的决定，项目发起人发生变更。听闻该新项目发起人的消息后，项目经理应该做什么？

A complex project with many stakeholders is in the execution phase, at which point the project sponsor changes due to management's decision. What should the project manager do when he hears about the new project sponsor?

A：将项目发起人变更的情况告知变更控制委员会(CCB)，并使之获批。

Inform the Change Control Board (CCB) of the change of the project sponsor and approve it

B：安排与新项目发起人举行紧急会议，告知其最新的工作绩效状态。

Schedule a contingency meeting with the new project sponsor to inform them of their latest work performance status

C：告知新项目发起人，需要先完成详细的工作交接才能推荐项目

Inform the new project sponsor that a detailed handover of work is required before the project can be recommended

D：安排与新项目发起人开会，了解战略目标，确保理解一致

Arrange meetings with new project sponsors to understand strategic objectives and ensure consistent understanding

正确答案：D

你的答案：A

解析：解析：PMBOK（6）P512-13.1.2.3，相关方分析。题干中的关键字：“项目发起人发生变更”。发起人发生了变化，由于发起人负责项目章程，所以需要和其确认他对项目目标的认识和当前项目目标保持一致，选项D正确。发起人变化和变更无关，排除A。了解工作绩效在D选项之后，排除B。选项C和题目无关，排除。

17、 [单选] 项目经理注意到，随着项目一拖再拖，很多相关方包括项目发起人都不出席项目状态通报会或阅读项目状态沟通报告。项目经理如何才能提高相关方的参与度？

The project manager notices that as the project drags on, and many stakeholders, including project sponsors, do not attend the project status briefing or read the project status communication report. How can a project manager increase the engagement of stakeholders?

A：修订沟通管理计划

Revise the communications management plan

B：使用他们自己的网络联系和影响相关方

Use their own networks to connect and influence stakeholders

C：修订相关方参与计划

Revise the stakeholder engagement plan

D：规定必须出席这些会议

Stipulate that these meetings must be attended

正确答案：C

你的答案：A

解析：解析：PMBOK（6）P535-13.4.3.3，相关方参与计划。题干中的关键字：“很多相关方不出席项目状态通报”“提高相关方参与度”。问题很明确，提高相关方的参与度，那么就是相关方参与计划相关的内容，选项C正确。沟通管理计划不提高相关方参与度，排除A。影响相关方也需要按照相关方参与计划来执行，排除B。规定必须出席太强硬，排除D

19、 [单选] 某项目经理正在制定最终的项目进度计划，正在与相关方一起进行最终审核。某技术相关方未出席该重要会议。项目经理接下来应该采取什么措施？

A project manager is developing a final project schedule and is working with the stakeholders to finally review it. A technique stakeholder did not attend the important meeting. What should the project supervisor do next?

A：将该相关方问题上报给项目发起人，以寻求支持

Escalate the stakeholder issue to the project sponsor for support

B：在该技术相关方参加的下次项目团队会议上讨论该技术问题

Discuss the technical issue at the next project team meeting attended by the technique stakeholder

C：提交项目进度计划，以便所有相关方都能达成共识并批准该进度计划

Submit a project schedule so that all stakeholders can reach agreements and approve the schedule

D：与该技术相关方开会，单独询问其意见，与其达成一致意见

Have a meeting with the technique stakeholders, ask for their opinions individually, and reach agreements with them

正确答案：D

你的答案：B

解析：解析：PMBOK（6）P535-13.4.2.6，会议。题干关键字：“相关方未出席该重要会议”。项目经理需要相关方确认该相关方对与目前项目决议的认同。选项D正确。项目未失控，无需上报，排除A。下次会议太消极，排除B。要先确认未参与的相关方的意见，排除C。

20、 [单选] 某国际公司的管理委员会决定为人力资源部门实施一个新的IT平台。该项目已指派项目经理，并已开始识别相关方。项目经理该怎么做？

The management committee of an international company decided to implement a new IT platform for human resources. The project has assigned a project manager and has begun to identify the stakeholders. What should the project manager do?

A：制作一幅思维导图，以确保发现所有有关的相关方和职能部门

Create a Mind-Mapping to ensure that all stakeholder and functions are found

B：让管理委员会提供该项目相关方的姓名

Let the Management Committee provide the name of the stakeholder

C：将项目发起人和项目团队成员纳入相关方名单，然后举办启动研讨会，从而发现其他相关方。

Include project sponsors and project team members in the list of stakeholders, and then hold an initiation workshop to identify other stakeholders

D：安排与人力资源部门总监开会，并让他们提供一份受影响相关方的名单

Arrange a meeting with the Director of Human Resources and have them provide a list of affected stakeholders

正确答案：C

你的答案：A

解析：解析：PMBOK（6）P514-13.1.2.5，会议。题干中的关键字：“开始识别相关方”。根据题干中的关键字，寻找识别相关方的工具以及识别相关方的产出，选项C正确。工具无法确保发现所有相关方，太绝对，排除A。识别相关方不止于名单，还要对相关方进行分析，排除BD。


