
\chapter{采购}

\section{错题分析}

1、 [单选] 项目团队确定他们需要采购的材料规格、数量和所需的质量等级。接下来应该完成哪一项？

The project team determined the specifications, quantities and required quality levels of the materials we needed to procure. Which one should be done next?

A：采购管理计划

The procurement management plan

B：采购工作说明书（SOW）

The procurement Statement of Work (SOW)

C：供方选择标准

The source selection criteria

D：需求文件

The requirements Documentation

正确答案：B

你的答案：A

解析：解析：PMBOK（6）P477，12.1.3.4-采购工作说明书。工作说明书会充分详细地描述拟采购的产品、服务或成果，以便潜在卖方确定是否有能力提供此类产品、服务或成果。选项A，采购管理计划记录了采购过程中的各项活动。选项C，供方选择标准是对潜在供应商的评判标准。选项D，需求文件是对收集到的需求的汇总。

4、 [单选] 项目团队需要获得市场上提供的其他数据，以明确范围中的产品规格。在这种情况下应该使用哪种类型的投标文件？

The project team needs to obtain additional data available on the market to clarify the product specifications in the scope. What type of bid document should be used in this case?

A：建议邀请书(RFP)

The request for proposal (RFP)

B：报价邀请书（RFQ)

The Request for Quotation (RFQ)

C：信息邀请书（RFI）

The request for information (RFI)

D：解决方案邀请书（RFS)

The request for solution (RFS)

正确答案：C

你的答案：A

解析：解析：PMBOK（6）P477，12.1.3.3招标文件-信息邀请书。如果需要卖方提供关于拟采购货物和服务的更多信息，就使用信息邀请书。随后一般还会使用报价邀请书或建议邀请书。

6、 [单选] 正处于执行阶段的某项目收到了一个装配组件，生产该组件需要更多的原材料，超过了与供应商所立合同的规定。供应商要求为所使用的额外材料付款。项目经理该如何处理这种情况？

A project in the execution phase receives an assembly component that requires more raw materials to produce than is required under a contract with the vendor. The vendor requests payment for the additional materials used. How should the project manager handle this situation?

A：接受该额外成本，向供应商支付该成本

Accept the additional cost and pay the vendor the cost

B：接收这批材料，但将付款延迟至项目结束

Receive the materials, but defer payment until the end of the project

C：就降低该批额外材料的价格展开谈判

Negotiate to reduce the price of the additional material

D：拒绝该额外成本，让供应商承担该成本

Reject the additional cost and let the vendor bear the cost

正确答案：D

你的答案：C

解析：PMBOK P489 实施采购-输出-协议（合同）：道德题目，项目经理应该按照合同做事，超出合同的部分不应该谈判或接受，D选项正确。干扰选项C，在实际情况中更常见，但不适用于PMI考试答题；



14、 [单选] 某组织希望为其网站门户中某—版块采用一种新的理念，CEO指定了项目经理，并希望立刻选定供应商。项目经理首先应该做什么？

An organization that wanted to adopt a new concept for a section of its web portal, the Chief Executive Officer designated a project manager and wanted to pick a vendor immediately. What should the project manager do first?

A：了解供应商的期望

Understand vendor expectations

B：为供应商设定资源选择原则

Set resource selection principles for vendors

C：根据组织过程资产创建一份供应商清单

Create a list of vendors based on the organizational process assets

D：为供应商制定行业标杆

Set industry baselines for vendors

正确答案：B

你的答案：C

解析：PMBOK P478 12.1.3.5 规划采购管理，输出，供方选择标准。关键词“采用新理念，选定供应商”意味着这种新理念是选择供应商的关键标准，B选项合适；

15、 [单选] 一家公司决定启动一个战略性项目，这项目将提供有竞争力的技术进步，该项目需要某 个特定性技术的专家，但组织无法提供准确的工作说明书（SOW)。 应该使用什么类型的合同？

A company decides to start a strategic project that will provide competitive technological advancement. The project requires experts in a specific technology, but the organization cannot provide an accurate statement of work (SOW). What type of contract should be used?

A：工料合同(T\&M)

Time and Material (T\&M) Contract

B：固定总价合同（FFP）

Firm Fixed Price (FFP) Contract

C：成本加固定费用合同（CPFF）

Cost plus Fixed Fee (CPFF) Contract

D：成本加激励费用合同(CPIF)

Cost plus Incentive Fee (CPIF) Contract

正确答案：A

你的答案：D

解析：PMBOK P472 工料合同定义：无法快速编制出准确的工作说明书，往往使用于聘用专家或外部资源。关键词“需要某 个特定性技术的专家，但组织无法提供准确的工作说明书”；选项A正确。

18、 [单选] 对于一个长期的基础设施项目，由于全球商业环境，项目经理必须获得风险最小的供应商。项目经理应该使用什么合同类型？

For a long-term infrastructure project, the project manager must acquire vendors with the lowest possible risk due to the global business climate. What type of contract should the project manager use?

A：固定总价合同（FFP）

Firm Fixed Price (FFP) Contract

B：成本加固定费用合同（CPFF）

Cost Plus Fixed Fee (CPFF) Contract

C：总价加激励费用合同（FPIF）

Fixed Price Incentive fee (FPIF) Contract

D：总价加经济价格调整合同（FPEPA)

Fixed Price with Economic Price Adjustment (FPEPA) Contract

正确答案：D

你的答案：A

解析：PMBOK P471, 12.1.1.6 合同类型，键字：长期，全球环境，风险最小。风险小需要采取总价合同，长期和不确定的全球环境需要加上经济价格调整，选项D正确。

21、 [单选] 采购部门要求项目经理提供一个项目的招标文件。这文件中应该包含哪些内容？

The procurement department asks the project manager for a project's bid documents. What should be included in these documents?

A：变更请求

The change requests

B：选择的卖方

The selected sellers

C：质量测量指标

The quality metrics

D：建议邀请书

The request for proposal

正确答案：D

你的答案：C

解析：PMBOK P477, 12.1.3.3招标文件，三种文件类型(书上原文)。采购文件中不包含变更请求，排除A。选择的卖方在采购管理计划中，排除B，质量测量指标是规划质量管理的输出。排除C。

25、 [单选] 某项目从一个新供应商订购了大量计算机备件。项目经理如何在达到项目时间表的同时确保产品质量？

A project ordered a large number of computer spare parts from a new vendor. How can the project manager ensure product quality while sticking to the project schedule?

A：确保提供商在发货前执行质量检查

Ensure that the provider performs a quality check before shipping

B：咨询主题专家(SME)应采用哪些标准

Request the subject matter expert (SME) what criteria should be used

C：根据故障率进行抽样检查

Carry out sample checks based on the failure rate

D：为客户购买更多备件

Buy more spare parts for the customers

正确答案：B

你的答案：C

解析：解析：PMBOK（6）P477，12.1.3.4-采购工作说明书。工作说明书会充分详细地描述拟采购的产品、服务或成果，以便潜在卖方确定是否有能力提供此类产品、服务或成果。情景的描述突出强调了如何满足质量要求，关于供应商交付产品、服务的要求要在采购工作说明书中明确。选项AC，无规矩不成方圆，质量检查要基于明确的质量标准，在选项B之后。选项D，按照合同的约定，供应商要承担在维保期间损坏备件的修复，所以为客户购买更多的备件是错误的。

26、 [单选] 在交付服务过程中某外包项目报告存在几个人为错误，包括未完成程序，未得到正确遵循的程序以及未执行的一般性检查，总经理发现了这些问题，要求尽快提高服务质量。项目经理首先应该做什么？

During the delivery of services, an outsourcing project reported several human errors, including uncompleted procedures, procedures that were not correctly followed, and general inspections that were not performed. The general manager discovered these issues and requested that the service quality be improved as soon as possible. What should the project manager do first?

A：将惩罚条款纳入合同，以迫使该外包公司提高服务交付人员的水平

Include penalty clauses in the contract to force the outsourcing company to improve the level of service delivery personnel

B：提交变更请求，将资源的技能质量标准纳入与该外包公司签订的合同

Submit a change request to include the skill quality criteria of the resource in the contract with the outsourcing company

C：要求该外包公司制定相关过程并对服务交付人员进行培训

Request the outsourcing company is d to develop processes and train service delivery personnel

D：如果外包公司未提高服务的交付质量，则与该公司终止合同

Terminate the contract with the company, if the outsourcing company fails to improve the quality of service delivery

正确答案：B

你的答案：C

解析：解析：PMBOK（6）P492，12.3-控制采购。“尽快”、“首先”应该做什么，处理情景中问题的顺序是：1.首先查阅合同中是否对此进行约定->2.如未约定，对合同进行变更，将检查标准纳入合同范围->3.根据新增加的合同条款监督合同的执行。按照这个顺序，选项B最靠前，正确。选项A，仅靠处罚不解决问题。选项C，所要求的相关服务要纳入合同范围，否则外包公司无履行义务。选项D，属于变更合同条款后的监督执行过程。



