
\chapter{风险}

\section{错题分析}

4、 [单选] 某软件开发项目的项目发起人对于整体负面项目风险可能会导致项目目标无法实现一事表示忧虑。项目经理可以采取以下哪种策略来调整整体项目风险的级别并优化实现项目目标的可能性？

The project sponsor of a software development project are concerned that overall negative project risks may result in the failure of the project objectives. Which of the following strategies can the project manager adopt to adjust the overall project risk level and optimize the likelihood of achieving project objectives?

A：举行风险评估研讨会，以便最大程度地降低风险影响估算

Execute risk assessment workshops to minimize risk impact estimation

B：将高收益要素增加到项目范围之中

Add high-yield elements to the project scope

C：将风险转移给第三方

Transfer risks to a third party

D：更改项目的范围和界限

Change the project scope and boundary

正确答案：D

你的答案：A

解析：解析：PMBOK（6）P445-11.5.2.7整体项目风险应对策略。取消项目范围中的高风险工作，是一种整个项目层面的规避措施。选项A，掩耳盗铃，降低估算，并不意味着风险不会发生。选项B，情景中没有提到高收益因素。选项C，注意情景中关注的是”调整整体项目风险的级别”，转移后，风险仍然存在，只是损失由第三方承担。


5、 [单选] 在审查项目报告时，项目经理识别到一个新的不稳定事件，可能延迟项目进度计划，为制定减少该事件的计划，项目经理首先应该做什么？

When reviewing the project report, the project manager identifies a new unstable matter that may delay the project schedule. What should the project manager do first in order to develop a plan to reduce this event?

A：更新风险登记册

Update the risk register

B：修订进度计划

Modify the schedule

C：评估对项目的整体影响

Assess the overall impact on the project

D：执行根本原因分析

Execute root cause analysis

正确答案：A

你的答案：C

解析：解析：PMBOK（6）P409，11.2-识别风险。“可能影响项目进度计划”，属于风险，要将其添加至风险登记册。接下来对风险进行评估，根据评估的结果制定风险应对计划，选项BCD在A选项之后。

7、 [单选] 一位经验丰富的项目团队成员通知项目经理，一个新风险未在风险登记册中列出。项目经理首先应该做什么？

An experienced project team member informs the project manager that a new risk is not listed in the risk register. What should the project manager do first?

A：与整个团队开会，讨论项目可能遇到的其他潜在风险

Have a meeting with the entire team to discuss other potential risks to the project

B：更新风险登记册

Update the risk register

C：告诉成员继续进行项目工作，因为风险可能不会对项目产生重要影响

Ask members to continue working because the risk may not have a significant impact on the project

D：分析风险

Analyze the risk

正确答案：B

你的答案：A

解析：解析：PMBOK（6）P453，11.7-监督风险。“新风险”直接未识别出，要首先更新到风险登记册，然后针对风险进行评估，根据评估的结果制定风险应对计划，选项D在B选项之后。选项A，风险已经被识别。选项C，未经评估即下定论，过于草率。

10、 [单选] 关键路径上的其中一项任务被认为具有高风险，因为它是一个全新的创新性任务。项目经理该做什么？

One of the tasks on the critical path is considered high-risk because it is a completely new and innovative task. What should the project manager do?

A：阐明特定需求，以降低不确定性

Clarify specific requirements to reduce uncertainty

B：更新质量管理计划，将更多测试包含进来

Update the quality management plan to include more tests

C：忽视该风险，因为有充足的应急储备

Ignore this risk because there are sufficient contingency reserves

D：更新进度计划，将该任务从关键路径移除

Update the schedule to remove the task from the critical path

正确答案：A

你的答案：B

解析：解析：PMBOK（6）P409，11.2-识别风险。“具有高风险”，要对风险进行分析评估，根据评估制定应对计划。选项BD是可能做出的应对计划。选项C是可能选择的应对策略。


17、 [单选] 在某道路基础设施项目施工过程中，已根据官方气象预报针对天气情况方面的风险估算安排了部分应急储备。但与项目规划过程中使用的最初预测相比，最新气象预报预测接下来数周的气象条件将更加恶劣。项目经理该做什么？

During the construction of a road infrastructure project, some contingency reserves have been arranged according to the official weather forecast's risk estimates for weather conditions. However, the latest weather forecast predicts worse weather conditions in the coming weeks than the initial forecasts used in the project planning process. What should the project manager do?

A：分析风险相对较小的项目活动剩余未用的应急储备，并在必要时动用它们

Analysis of unused contingency reserves for relatively low-risk project activities and, where necessary, their use

B：由于根据气象预报可能会出现额外风险，向变更控制委员会(CCB)提出动用管理储备的请求

Submit a request to the Change Control Board (CCB) to use the management reserve, as additional risks may arise based on weather forecasts

C：确定剩余的应急储备是都足以应对不确定的气象条件，并在必要时提出变更请求

Determine that the remaining contingency reserves are sufficient to cope with uncertain weather conditions and, if necessary, propose change requests

D：检查剩余的项目预算是否足以涵盖更加恶劣的气象条件可能会造成的额外风险

Check that the remaining project budget is sufficient to cover the additional risks that may result from harsher weather conditions

正确答案：C

你的答案：D

解析：PMBOKP456 11.7.2.1-监督风险-工具技术-储备分析。关键词“最新预测，更加恶劣”，更加恶劣的情况即风险的危害加强，那么之前规划的风险储备可能是不充足的，需要进行补充，补充储备是需要变更的，所以选项C正确。应急储备专项专用，未用到的需要通过变更释放资源，排除A。管理储备适用于未知未知风险，此风险已知，排除B。D描述正确但是没有C全面，排除D。


18、 [单选] 项目经理向某部门交付系统。其中一个项目供应商在交付关键系统组件方面总是延迟。该项目非常重要，不得落后于所承诺的进度。项目经理应如何避免将来发生这种情况？

The project manager delivers the system to a department. One of the project vendors is always delayed in delivering critical system components. The project was too important to fall behind the promised schedule. What should the project manager do to avoid this happening in the future?

A：通过聘请有着出色交付业绩记录的另一个供应商避免风险

Avoid risks by hiring another vendor with an excellent delivery track record

B：将风险上报给项目发起人，因为风险超出了项目控制

Escalate the risk to the project sponsor because the risk is beyond the control of the project

C：通过与现任供应商商定经济处罚条款减轻风险

Mitigate risk by agreeing on economic penalties with current vendors

D：通过购买针对现任供应商迟延交付的保险来转移风险

Transfer risk by purchasing insurance against delayed delivery by current vendors

正确答案：A

你的答案：D

解析：PMBOKP443-11.5.2.4-规划风险应对-工具技术-威胁应对策略-规避。题目关键字：“如何避免”避免指风险不在发生，选择规避策略，选项A正确。上报，减轻，转移都无法避免风险的发生，排除BCD。

19、 [单选] 在某项目执行阶段，项目经理获悉，某项目需求可能会发生变更，这可能会对该项目的成本和进度产生重大影响。在这种情况下，项目经理应该做什么？

During a project execution phase, the project manager is informed that the requirements for a project may change, which can have a significant impact on the cost and schedule of the project. What should the project manager do in this case?

A：将对该需求的潜在变更增加到项目章程中，然后立即调整项目成本和进度计划文件

Add potential changes to the requirement to the project charter, and then immediately adjust the project cost and schedule documentation

B：对风险登记册不作修改，但就对项目成本和进度计划的影响征得项目发起人批准

There are no changes to the risk register, but the impact on project cost and schedule is approved by the project sponsor

C：分配必要人员应对该潜在变更，尽管该应对措施未在风险应对计划中获批

Assign the necessary personnel to respond to this potential change, although the response was not approved in the risk response plan

D：提出针对潜在变更的变更请求，并征得变更控制委员会（CCB）批准

Make a change request for a potential change and obtain approval from the Change Control Board (CCB)

正确答案：C

你的答案：D

解析：PMBOK 11.6.3.2 实施风险应对-输出-项目团队派工单；关键词“执行阶段，PM获悉，可能变更，重大影响”表明这是一个风险，且已被识别，应当规划并实施应对，其中实施应对的一个关键措施就是分配资源以应对风险。项目章程变更意味着项目的重新授权，排除A，从题目中很明显发现了新的风险，需要更新风险登记册，排除B。变更还没有发生就直接申请潜在变更不正确，排除D。



23、 [单选] 项目技术团队审核期间，先前识别的一个风险现已通过审核。根据这些信息，项目经理应采取哪些措施？

During the project technical team review, a risk previously identified has now been reviewed. Based on this information, what actions should the project manager take?

A：在风险登记册中关闭该风险，但要将相关的应急储备作为未来风险的缓冲

Close risks in the risk register, but use the relevant contingency reserves as a buffer against future risks

B：继续监督该风险，直到项目完成，因为未来还可能会发生该风险

Continue to monitor the risk until the project is completed, as the risk may occur in the future

C：利用相关的风险储备分配额外资源，从而加速项目的推进

Use associated risk reserves to allocate additional resources to accelerate project schedule

D：在风险登记册中将该风险标记为已关闭，并减少相关的应急储备

Mark the risk as closed in the risk register and reduce the relevant contingency reserves

正确答案：D

你的答案：B

解析：PMBOK 11.7.3.4 P458 监督风险，输出，项目文件更新，风险登记册。关键词“审核期间，识别的风险已通过审核”表明风险应对已实施，可以关闭，并将结果更新到风险登记册上。该题容易误导的选项是B，确实应该继续监督该风险直到项目完成，但不是因为该风险还会发生，而是因为需要监督风险是否发生变化而导致之前实施的应对不再起作用了；由于风险关闭而释放的应急储备应该释放给组织，而不是在项目内继续使用，排除AC，已关闭的风险不用再继续监督。若有新风险或此生风险应该再记录在登记册中。

