
\chapter{RNN}
\label{cha:rnn}

RNN stands for recurrent neural network.
RNN is neural network that has recurrent layers.
While CNNs can efficiently process spatial information, RNNs are designed to better handle sequential information.
RNNs introduce state variables to store past information, together with the current inputs, to determine the current outputs.

\section{Recurrent}
\label{sec:recurrent}

\begin{equation}
  \label{eq:3}
  s^{(t)} = f(s^{(t-1)}; \theta)
\end{equation}
Where \(s^{(t)}\) is the state or output at time \(t\), \(f\) is the function, \(\theta\) is the parameter.
Equation \ref{eq:3} is recurrent because \(s\) at time \(t\) refer to itself at time \(t-1\).


\begin{equation}
  \label{eq:4}
  s^{(t)} = f(s^{(t-1)}, x^{(t)}; \theta)
\end{equation}
This equation is similar to equation \ref{eq:3} but with input \(x^{(t)}\).


In natural language processing (NLP), we usually need to compute
\begin{equation}
  \label{eq:nlp}
  P(x_{t}|x_{t-1},\ldots,x_{1})
\end{equation}

With n-gram model (Section \ref{sec:n-grams}), the number of model parameters increase exponentially as n increase.
We need to store \(|V|^{n}\) numbers for a vocabulary set \(V\).
In this case, we usually use a latent variable model:
\begin{equation}
  \label{eq:rnn-model}
  P(x_{t}|x_{t-1},\ldots,x_{1}) \approx P(x_{t}|h_{t-1})
\end{equation}
and
\begin{equation}
  \label{eq:latent}
  h_{t} = f(x_{t},h_{t-1})
\end{equation}


\section{Properties}
\label{sec:properties}


RNN leverages three important ideas:
\begin{itemize}
\item sparse interaction.
\item parameter sharing.
\item 
\end{itemize}

\subsection{Sparse interaction}
\label{sec:sparse-interaction}


Comparing to traditional neural network, it has sparse interaction.
For example, in Equation \ref{eq:4}, \(s^{(t)}\) is determined by \(s^{(t-1)}\) and \(x^{(t)}\).
It has no direct interaction with \(x^{(1)}, x^{(2)}, \ldots x^{(t-1)}\) because they are contained in \(s^{(t-1)}\).


\subsection{Parameter sharing}
\label{sec:parameter-sharing}

In recurrent layers, the same parameters defined in function \(f\) are used at every position of the input.






%%% Local Variables:
%%% mode: latex
%%% TeX-master: "deep-learning"
%%% End:
